---
title: "2020 Stack Overflow Developer Survey Results"
output:
  html_document:
    df_print: paged
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

The Stack Overflow Developer Survey 2020 was applied from February 5th to the 28th. 
There were a total of 64461 respondents for a total of 61 questions. However, less than 30 questions are considered for this analysis.

The main objective of this analysis is to explore who are the respondents, such as how they use code professionally, their country, academic background, technologies they work with, etc.

The original dataset can be downloaded from Stack Overflow's website [here](https://insights.stackoverflow.com/survey).

# Imports
```{r}
require(RPostgreSQL)
require(ini)
library(plotly)
```

# Data Analysis

## Connect to Database


```{r}
# Load host, database, user and password information stored in the "postgresql" group
db_info <- read.ini("db.ini")$postgresql
```


```{r}
db_driver <- dbDriver("PostgreSQL")
db_connection <- dbConnect(
    db_driver, 
    dbname = db_info$database, 
    host = db_info$host, 
    user = db_info$user, 
    password = db_info$password
)
```

## General Data

### Gender of Respondents


```{r}
query_str <- "
    SELECT gender, COUNT(gender) AS gender_count
    FROM respondents
    WHERE gender IS NOT NULL
    GROUP BY gender;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_pie(
        data = df, 
        labels = df$gender, 
        values = df$gender_count
    ) %>% 
    layout(title = "Gender of Respondents")
fig
```

### Age of Respondents


```{r}
query_str <- "
    SELECT age, COUNT(age) AS age_count
    FROM respondents
    WHERE age BETWEEN 10 AND 80
    GROUP BY age;
"
df <- dbGetQuery(db_connection, query_str)

# Relative frequencies for the key takeaways (single-choice question)
age_20_30 <- sum(df[df$age >= 20 & df$age <=30,"age_count"] / sum(df$age_count))

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$age, 
        y = df$age_count
    ) %>% 
    layout(title = "Age of Respondents")
fig
```

### Country of Respondents


```{r}
query_str <- "
    SELECT country, COUNT(country) AS country_count
    FROM respondents
    GROUP BY country
    ORDER BY country_count DESC
    LIMIT 15;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$country_count),]

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$country_count,
        y = df$country
    ) %>%
    layout(
        title = "Country of Respondents",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$country_count
        )
    )
fig
```

### Education Level of Respondents


```{r}
query_str <- "
    SELECT ed_level, COUNT(ed_level) AS ed_level_count
    FROM respondents
    WHERE ed_level IS NOT NULL
    GROUP BY ed_level
    ORDER BY ed_level_count DESC;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$ed_level_count),]

# Relative frequencies for the key takeaways (single-choice question)
ed_level_bachelor <- df[df$ed_level == "Bachelor's degree","ed_level_count"] / sum(df$ed_level_count)
ed_level_masters <- df[df$ed_level == "Master's degree","ed_level_count"] / sum(df$ed_level_count)

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$ed_level_count,
        y = df$ed_level
    ) %>%
    layout(
        title = "Education Level of Respondents",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$ed_level_count
        )
    )
fig
```

### Hobbyist Developers


```{r}
query_str <- "
    SELECT hobbyist, COUNT(hobbyist) AS hobbyist_count
    FROM respondents
    WHERE hobbyist IS NOT NULL
    GROUP BY hobbyist;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_pie(
        data = df, 
        labels = df$hobbyist, 
        values = df$hobbyist_count
    ) %>% 
    layout(title = "Do you code as a hobby?")
fig
```

### Key Takeaways

* 9 out of every 10 respondents identify themselves as men
* Approximately 52.14% of the respondents are between the ages of 20 and 30
* Twelve thousand respondents are from the United States of America, followed by India with eight thousand and the United Kingdom with almost four thousand respondents
* Bachelor and Master's Degrees are the most common education levels, representing 46.22% and 22.83% of the answers, respectively
* Almost 80% of the respondents reported coding as a hobby

## Technology-related Data

### Programming Languages


```{r}
query_str <- "
    SELECT prog_language, COUNT(prog_language) AS prog_language_count
    FROM prog_languages
    GROUP BY prog_language
    ORDER BY prog_language_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$prog_language, 
        y = df$prog_language_count
    ) %>% 
    layout(
        title = "10 Most Used Programming Languages",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$prog_language_count
        )
    )
fig
```

### Databases


```{r}
query_str <- "
    SELECT prog_language, COUNT(prog_language) AS prog_language_count
    FROM prog_languages
    GROUP BY prog_language
    ORDER BY prog_language_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$prog_language, 
        y = df$prog_language_count
    ) %>% 
    layout(
        title = "10 Most Used Programming Languages",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$prog_language_count
        )
    )
fig
```

### Web Frameworks


```{r}
query_str <- "
    SELECT web_fw, COUNT(web_fw) AS web_fw_count
    FROM web_fws
    GROUP BY web_fw
    ORDER BY web_fw_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$web_fw, 
        y = df$web_fw_count
    ) %>% 
    layout(
        title = "10 Most Used Web Frameworks",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$web_fw_count
        )
    )
fig
```

### Collaboration Tools


```{r}
query_str <- "
    SELECT collab_tool, COUNT(collab_tool) AS collab_tool_count
    FROM collab_tools
    GROUP BY collab_tool
    ORDER BY collab_tool_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$collab_tool, 
        y = df$collab_tool_count
    ) %>% 
    layout(
        title = "10 Most Used Collaboration Tools",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$collab_tool_count
        )
    )
fig
```

### Platforms


```{r}
query_str <- "
    SELECT platform, COUNT(platform) AS platform_count
    FROM platforms
    WHERE platform IS NOT NULL
    GROUP BY platform
    ORDER BY platform_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$platform, 
        y = df$platform_count
    ) %>% 
    layout(
        title = "10 Most Used Platforms",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$platform_count
        )
    )
fig
```

### Operating Systems


```{r}
query_str <- "
    SELECT op_sys, COUNT(op_sys) AS op_sys_count
    FROM respondents
    WHERE op_sys IS NOT NULL
    GROUP BY op_sys
    ORDER BY op_sys_count DESC;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$op_sys, 
        y = df$op_sys_count
    ) %>% 
    layout(
        title = "Most Used Operating Systems",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$op_sys_count
        )
    )
fig
```

### Key Takeaways

* JavaScript, HTML/CSS and SQL are the most used (programming) languages
* The most used database engines are MySQL, PostgreSQL and Microsoft SQL Server
* jQuery, React.js and Angular are the three most used web development frameworks
* GitHub, Slack and Jira were reported as the three most used collaboration tools
* Linux, Windows and Docker are the three leading development platforms
* 48.32% of the respondents use Windows as their operating system, 26.56% use Linux-based systems and 25.03% use MacOS

## Professional Life Data

### Employment Status


```{r}
query_str <- "
    SELECT employment_status, COUNT(employment_status) AS employment_status_count
    FROM respondents
    WHERE employment_status IS NOT NULL
    GROUP BY employment_status
    ORDER BY employment_status_count DESC;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$employment_status_count),]

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$employment_status_count,
        y = df$employment_status
    ) %>%
    layout(
        title = "Employment Status of Respondents",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$employment_status_count
        )
    )
fig
```

### Most Common Developer Roles


```{r}
query_str <- "
    SELECT COUNT(DISTINCT id) AS num_respondents
    FROM respondents;
"
num_respondents <- dbGetQuery(db_connection, query_str)[1, "num_respondents"]
```


```{r}
query_str <- "
    SELECT dev_role, COUNT(dev_role) AS dev_role_count
    FROM dev_roles
    WHERE dev_role IS NOT NULL
    GROUP BY dev_role
    ORDER BY dev_role_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$dev_role_count),]

# Relative frequencies for the key takeaways
dev_role_backend <- df[df$dev_role == "Back-end developer","dev_role_count"] / num_respondents
dev_role_fullstack <- df[df$dev_role == "Full-stack developer","dev_role_count"] / num_respondents
dev_role_frontend <- df[df$dev_role == "Front-end developer","dev_role_count"] / num_respondents

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$dev_role_count,
        y = df$dev_role
    ) %>%
    layout(
        title = "10 Most Common Developer Roles",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$dev_role_count
        )
    )
fig
```

### Annual Compensation


```{r}
query_str <- "
    SELECT yearly_compensation
    FROM respondents
    WHERE yearly_compensation BETWEEN 0 AND 200000;
"
df <- dbGetQuery(db_connection, query_str)
# round(df$yearly_compensation / length(df$yearly_compensation), 2)

bins <- list(start = 0, end = 210000, size = 15000)

fig <- plot_ly() %>%
    add_histogram(
        data = df,
        x = df$yearly_compensation,
        xbins = bins
    ) %>%
    layout(
        title = "Annual Compensation"
    )
fig
```

### Average Annual Compensation by Programming Language


```{r}
query_str <- "
    SELECT 
        prog_languages.prog_language AS prog_language, 
        AVG(respondents.yearly_compensation) AS compensation_avg
    FROM prog_languages
    INNER JOIN respondents
    ON prog_languages.respondent_id = respondents.id
    WHERE yearly_compensation BETWEEN 0 AND 200000
    GROUP BY prog_language
    ORDER BY compensation_avg DESC
    LIMIT 20;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$compensation_avg),]

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$compensation_avg,
        y = df$prog_language
    ) %>%
    layout(
        title = "Average Annual Compensation by Programming Language (Top 20)",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$compensation_avg
        )
    )
fig
```

### Average Annual Compensation by Developer Role


```{r}
query_str <- "
    SELECT 
        dev_roles.dev_role AS dev_role, 
        AVG(respondents.yearly_compensation) AS compensation_avg
    FROM dev_roles
    INNER JOIN respondents
    ON dev_roles.respondent_id = respondents.id
    WHERE 
        yearly_compensation BETWEEN 0 AND 200000
        AND dev_role IS NOT NULL
    GROUP BY dev_role
    ORDER BY compensation_avg DESC;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$compensation_avg),]

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$compensation_avg,
        y = df$dev_role
    ) %>%
    layout(
        title = "Average Annual Compensation by Developer Role (Top 20)",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$compensation_avg
        )
    )
fig
```

### Key Takeaways

* 70.9% of respondents are employed full-time. The second most common employment status is student, for 12.2% of the respondents
* Back-end, full-stack and front-end are the most common developers roles, with 42.24%, 42.08% and 28.38% of the developers, respectively, having one or more of these roles
* The annual income of respondents is concentrated in the interval between 30k USD and 105k USD
* Scala, Ruby and Go developers are the most well-paid, all having an annual income above 90k USD
* Engineering Managers are the most well-paid developers, with an annual income of over 100k USD. Senior Executives come in second with 98k USD and Site Reliability Engineers in third with 98k USD as well

## Other Data

### Hours of Work per Week


```{r}
query_str <- "
    SELECT work_week_hrs
    FROM respondents
    WHERE work_week_hrs BETWEEN 0 AND 100;
"
df <- dbGetQuery(db_connection, query_str)
# round(df$yearly_compensation / length(df$yearly_compensation), 2)

bins <- list(start = 0, end = 100, size = 10)

fig <- plot_ly() %>%
    add_histogram(
        data = df,
        x = df$work_week_hrs,
        xbins = bins
    ) %>%
    layout(
        title = "Hours of Work per Week"
    )
fig
```

### Overtime Work Frequency


```{r}
query_str <- "
    SELECT overtime, COUNT(overtime) AS overtime_count
    FROM respondents
    WHERE overtime IS NOT NULL
    GROUP BY overtime;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_pie(
        data = df, 
        labels = df$overtime, 
        values = df$overtime_count
    ) %>% 
    layout(title = "Overtime Work Frequency")
fig
```

### Undergrad Major of Respondents


```{r}
query_str <- "
    SELECT undegrad_major, COUNT(undegrad_major) AS undegrad_major_count
    FROM respondents
    WHERE undegrad_major IS NOT NULL
    GROUP BY undegrad_major
    ORDER BY undegrad_major_count DESC;
"
df <- dbGetQuery(db_connection, query_str)
df <- df[order(df$undegrad_major_count),]

fig <- plot_ly() %>%
    add_bars(
        data = df,
        x = df$undegrad_major_count,
        y = df$undegrad_major
    ) %>%
    layout(
        title = "Undergrad Major of Respondents",
        yaxis = list(
            categoryorder = "array",
            categoryarray = df$undegrad_major_count
        )
    )
fig
```

### Solutions Used When Stuck


```{r}
query_str <- "
    SELECT stuck_solution, COUNT(stuck_solution) AS stuck_solution_count
    FROM stuck_solutions
    WHERE stuck_solution IS NOT NULL
    GROUP BY stuck_solution
    ORDER BY stuck_solution_count DESC;
"
df <- dbGetQuery(db_connection, query_str)

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$stuck_solution, 
        y = df$stuck_solution_count
    ) %>% 
    layout(
        title = "Solutions Used When Stuck",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$stuck_solution
        )
    )
fig
```

### Most Important Job Factors


```{r}
query_str <- "
    SELECT COUNT(DISTINCT id) AS num_respondents
    FROM respondents;
"
num_respondents <- dbGetQuery(db_connection, query_str)[1, "num_respondents"]
```


```{r}
query_str <- "
    SELECT job_factor, COUNT(job_factor) AS job_factor_count
    FROM job_factors
    GROUP BY job_factor
    ORDER BY job_factor_count DESC
    LIMIT 10;
"
df <- dbGetQuery(db_connection, query_str)

# Relative frequencies for the key takeaways
job_factor_tech <- df[df$job_factor == "Technical compatibility","job_factor_count"] / num_respondents
job_factor_culture <- df[df$job_factor == "Company culture","job_factor_count"] / num_respondents
job_factor_schedule <- df[df$job_factor == "Flexible schedule","job_factor_count"] / num_respondents

fig <- plot_ly() %>% 
    add_bars(
        data = df, 
        x = df$job_factor, 
        y = df$job_factor_count
    ) %>% 
    layout(
        title = "10 Most Important Job Factors",
        xaxis = list(
            categoryorder = "array",
            categoryarray = df$job_factor_count
        )
    )
fig
```

### Key Takeaways

* The large majority of respondents spend 40 to 50 hours at work, each week
* A quarter of the respondents report working overtime often, that is, at least once per week. On the other hand, a quarter of the respondents also report never working overtime or rarely
* Computer Science, Computer Engineering and Software Engineering are by far the most common undergrad major. Engineering as an area of study comes in second, followed by Information Systems/Information Technologies and System Administration
* Visiting Stack Overflow is the most used solution when stuck on a technical problem.The second most used solution is doing other work and returning to the problem later, and the third is watching help and/or tutorial videos
* Knowing people can take more than one factor into consideration when choosing a job, technical compatibility is the most sough after, by 39.3% of the respondents. Company culture comes second with 34.1% and flexible schedule third with 33.6%

## Close Database Connections


```{r}
dbDisconnect(db_connection)
```
